---
title: "PFNA accumulating strains"
author: "Anna Lindell"
date: "August 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(tidyverse)
library(readr)
library(ggh4x)
library(ggplot2)
```




# Quality control

Internal standard:
```{r eval = FALSE}
data <- read.table(file = "data/assay3_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)

is <- data %>% filter(group != "Blank")

ggplot(is, aes(x = file, y = is.resp)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_blank()) +
  facet_wrap(~batch, scales = "free") +
  labs(y = "Internal Standard Response", x = "Data File") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/assay3_is.png",
       width = 20, height = 15, units = "cm")

cv.is.data <- print.data.frame(is %>% group_by(batch) %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), is.resp))

summary(cv.is.data$cv)
```


|          | Min   | Median | Mean  | Max    |
|----------|-------|--------|-------|--------|
| IS       | 5.2 % | 13.6 % | 16 %  | 40 %   |






Quality control:
```{r eval = FALSE}
qc <- data %>% filter(group == "QC")

ggplot(qc, aes(x = file, y = compound.resp)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_blank()) +
  facet_wrap(~batch, scales = "free") +
  labs(y = "Response", x = "Data File") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/assay3_qc_resp.png",
       width = 20, height = 15, units = "cm")

ggplot(qc, aes(x = file, y = ratio)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_blank()) +
  facet_wrap(~batch, scales = "free") +
  labs(y = "Ratio", x = "Data File") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/assay3_qc_ratio.png",
       width = 20, height = 15, units = "cm")

cv.qc.resp <- print.data.frame(qc %>% group_by(batch) %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), compound.resp))
cv.qc.ratio <- print.data.frame(qc %>% group_by(batch) %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), ratio))

summary(cv.qc.resp$cv)
summary(cv.qc.ratio$cv)

rm(list = ls())
```

|          | Min   | Median | Mean  | Max    |
|----------|-------|--------|-------|--------|
| Response | 1.3 % | 6.3 %  | 14 %  | 56 %   |
| Ratio    | 2.9 % | 8.7 %  | 14 %  | 31 %   |


QC's quite stabel over time. Slight signal drift for some compounds.

Using the Response or ratio does not make a big change for the QC. Concentration calculated based on compounds response.





# Median Reduction:

```{r eval = FALSE}
data <- read.table(file = "data/assay3_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data <- data %>% 
  filter(group %in% c("Compound control", "Supernatant", "Whole culture"))

cc_median <- data %>%
  filter(group == "Compound control") %>%
  group_by(batch) %>%
  summarise(median = median(conc)) 

n <- data %>%
  group_by(batch) %>%
  count()
n <- n$n

cc_median <- cc_median$median
cc_median <- rep(cc_median,  n)

median <- data %>%
  arrange(experiment, batch, compound, strain, group, replicate) %>%
  mutate(percent = conc / cc_median * 100)

write.table(median,
            file = "data/assay3_medianreduction.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```






### Plot Samples

Plot data:
```{r eval = FALSE}
data <- read.table(file = "data/assay3_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

bat <- unique(data$batch)

for (c in seq_along(bat)) {
  subset <- data %>% filter(batch == bat[c])
  control <- subset %>% filter(group == "Compound control")
  median <- median(control$conc)
  plot <- ggplot(subset, aes(x = group, y = conc)) +
    geom_hline(yintercept = median, color = "grey") +
    geom_jitter(width = 0.1, size = 3, shape = 1) +
    facet_grid(~strain.id, scales = "free_x") +
    theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          text = element_text(size=10),
          axis.text.x = element_text(size=10, angle = 90, , vjust = 0.5, hjust = 0.5, color = "black"),
          axis.text.y = element_text(size=10, color = "black"),
          axis.title.x = element_blank(),
          strip.background = element_blank()) +
    labs(y = "PFNA Recovery [uM]") +
    ylim(0,NA)
  ggsave(plot,
           file = paste("figures/batch", bat[c], "_conc.png", sep = ""),
           width = 40, height = 7, units = "cm")
}


ggplot(subset(data, data$group != "Compound control"), aes(x = group, y = percent)) +
  geom_hline(yintercept = 100, color = "grey") +
  geom_jitter(width = 0.1, size = 3, shape = 1) +
  facet_wrap(~strain.id) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 0.5, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        strip.background = element_blank()) +
  labs(y = "PFNA recovery [%]") +
  ylim(0,NA)
ggsave(file = "results/assay3_percent.png",
       width = 50, height =50, units = "cm")

rm(list = ls())
```





# Statistical analysis

T-test:
```{r eval=FALSE}
data <- read.table(file = "data/assay3_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

bugs <- subset(data, data$strain != "Compound control")
control <- subset(data, data$strain == "Compound control")

bat <- unique(bugs$batch)
strains <- unique(bugs$strain)
lev <- unique(bugs$group)

n <- 1
p_value <- data.frame(batch = numeric(),
                      strain = character(),
                      group = character(),
                      p.value = numeric())
for (c in seq_along(bat)) {
  bugs1 <- bugs %>% filter(batch == bat[c])
  control1 <- control %>% filter(batch == bat[c])
  for (i in seq_along(strains)) {
    bugs2 <- bugs1 %>% filter(strain == strains[i])
    if(nrow(bugs2) ==0) next
    for (l in seq_along(lev)) {
      bugs3 <- bugs2 %>% filter(group == lev[l])
      test <- t.test(x = bugs3$conc, y = control1$conc, alternative = "two.sided")
      p_value[n, 1:4] <- c(bat[c], strains[i], lev[l], test$p.value)
      n <- n + 1
    }
  }
}

p_value$p.value <- as.numeric(p_value$p.value)
data$batch <- as.character(data$batch)

adj_data <- p_value %>%
  mutate(p_adj = p.adjust(p.value, method = "fdr")) %>%
  arrange(batch, strain, group)

new.data <- data %>%
  group_by(experiment, batch, strain.id, strain, group, compound, unit) %>%
  summarise(conc = median(conc),
            percent = median(percent)) %>%
  full_join(adj_data, by = c("batch", "strain", "group"))

# take mean of same B. uniformis strains from batch 1 and 3
bu <- new.data %>% 
  filter(strain.id == "NT5002") %>%
  group_by(experiment, strain.id, strain, group, compound, unit) %>%
  summarize(conc = mean(conc),
            percent = mean(percent),
            p.value = mean(p.value),
            p_adj = mean(p_adj)) %>%
  add_column(batch = rep("1 & 3", 2))
new.data <- new.data %>%
  filter(strain.id != "NT5002") %>%
  rbind(bu)

# add hit column
new.data$significant <- ifelse(new.data$percent <= 80 & new.data$p_adj < 0.05, "yes", "no")

# save data file
write.table(new.data,
            file = "results/assay3_results.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```






# Results analysis:

Plot supernatant reduction in heatmap:
```{r eval = FALSE}
data <- read.table(file = "results/assay3_results.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data <- data %>%
  filter(group == "Supernatant") %>%
  mutate(percent.accumulation = 100 - percent)
data$percent.accumulation[data$percent.accumulation < 0] <- 0

ggplot(data, aes(x = group, y=fct_rev(strain))) +
  geom_tile(aes(fill = percent.accumulation), color = "black") +
  geom_text(aes(label = round(percent.accumulation, 0))) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black", face = "italic"),
        text = element_text(size = 10)) +
  scale_fill_gradient2(name = "PFNA accumulation [%]", high = "#6f89a2", low = "white", limits=c(0,100))
ggsave(file = "results/assay3_heatmap_detailed.png",
       width = 12, height = 50, units = "cm")

ggplot(data, aes(x = group, y=fct_rev(strain))) +
  geom_tile(aes(fill = percent.accumulation), color = "black") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black", face = "italic"),
        text = element_text(size = 10)) +
  scale_fill_gradient2(name = "PFNA accumulation [%]", high = "#6f89a2", low = "white", limits=c(0,100))
ggsave(file = "results/assay3_heatmap.png",
       width = 12, height = 50, units = "cm")

write.table(data,
            file = "results/assay3_heatmap.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```


Plot distributions:
```{r eval = FALSE}
data <- read.table(file = "results/assay3_heatmap.txt", 
                   sep = "\t", dec = ".", header = TRUE)

# plot total distribution
ggplot(data, aes(x = percent.accumulation)) +
  geom_density() +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        text = element_text(size = 10)) +
  labs(x = "PFNA accumulation [%]", y = "Density")
ggsave(file = "results/assay3_distribution.png",
       width = 10, height = 10, units = "cm")

# add strain data and save
strains <- read.table(file = "data/strains.txt", 
                   sep = "\t", dec = ".", header = TRUE)
data.strains <- merge(data, strains, by = c("strain", "strain.id"), all = TRUE)

write.table(data.strains,
            file = "results/assay3_distribution_gram_phylum.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

# plot gram strain distribution
ggplot(subset(data.strains, phylum != "Yeast"), aes(x = percent.accumulation, linetype = gram.stain)) +
  geom_density() +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 10, color = "black"),
        axis.title.y = element_text(size = 10, color = "black"),
        text = element_text(size = 10),
        legend.position = c(0.75, 0.8),
        legend.key.size = unit(0.6,"line"),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10)) +
  labs(x = "PFNA accumulation [%]", y = "Density", linetype = "Gram-stain")
ggsave(file = "results/assay3_distribution_gramstain.png",
       width = 10, height = 10, units = "cm")

ggplot(data.strains, aes(x = phylum, y = percent.accumulation)) +
  geom_violin(width = 1) +
  geom_jitter(aes(color = phylum), width = 0.1, size = 2, alpha = 0.5) +
  guides(fill="none") +
  scale_color_manual(values = c("#ee3f36", "#ee8300", "#8bc53f", "#f9ec31", "#ee00ee", "#00afee", "grey")) +
  scale_fill_manual(values = c("#ee3f36", "#ee8300", "#8bc53f", "#f9ec31", "#ee00ee", "#00afee", "grey")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black", angle = 45, hjust = 1),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10, color = "black"),
        text = element_text(size = 10),
        legend.position = "none") +
  labs(y = "PFNA accumulation [%]")
ggsave(file = "results/assay3_phylum_violin.pdf",
       width = 10, height = 8, units = "cm")

rm(list = ls())
```




Correlation between growth and resting screen:
```{r eval = FALSE}
data <- read.table(file = "results/correlation_growth_resting/assay2_3_correlation.txt", 
                   sep = "\t", dec = ".", header = TRUE)

cor.test(data$acc.growth, data$acc.rest, method = "pearson")
# r = 0.76; p = 0.002; strong positive correlation
cor.test(data$acc.growth, data$acc.rest, method = "spearman")
# rho = 0.64; p = 0.015; strong positive correlation

library(ggpubr)
ggplot(data, aes(x = acc.growth, y = acc.rest)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")), size = 3, label.x = 1, label.y = 50) +
  annotate(geom="text", label="r = 0.76; p = 0.002", x=17, y=47, size = 3) +
  annotate(geom="text", label="rho = 0.64; p = 0.015", x=17, y=44, size = 3) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 10, color = "black"),
        axis.title.y = element_text(size = 10, color = "black"),
        text = element_text(size = 10)) +
  labs(y = "PFNA accumulation in resting assay [%]", x = "PFNA accumulation in growth assay [%]")
ggsave(file = "results/correlation_growth_resting/assay3_correlation.png",
       width = 10, height = 10, units = "cm")


# Correlation between growth and resting screen for OD corrected accumulation:
ggplot(data, aes(x = norm.acc.growth, y = norm.acc.rest)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")), size = 3, label.x = 1, label.y = 15) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 10, color = "black"),
        axis.title.y = element_text(size = 10, color = "black"),
        text = element_text(size = 10)) +
  labs(y = "Normalised PFNA accumulation in resting assay [%]", x = "Normalised PFNA accumulation in growth assay [%]")
ggsave(file = "results/correlation_growth_resting/assay3_OD_corrected_correlation.png",
       width = 10, height = 10, units = "cm")

# Correlation between growth and resting screen for OD corrected accumulation after removing outlier:
data1 <- data %>% filter(id != "NT5021")

cor.test(data1$acc.growth, data1$acc.rest, method = "pearson")
# r = 0.73; p = 0.004; strong positive correlation
cor.test(data1$acc.growth, data1$acc.rest, method = "spearman")
# rho = 0.56; p = 0.049; strong positive correlation

ggplot(data, aes(x = norm.acc.growth, y = norm.acc.rest)) +
  geom_point() +
  geom_smooth(data = data1, method = "lm", se = F) +
  stat_regline_equation(data = data1, aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")), size = 3, label.x = 1, label.y = 15) +
  annotate(geom="text", label="r = 0.73; p = 0.004", x=17, y=14, size = 3) +
  annotate(geom="text", label="rho = 0.56; p = 0.049", x=17, y=13, size = 3) +
  annotate(geom="text", label="Outlier excluded", x=57, y=3.3, size = 3) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 10, color = "black"),
        axis.title.y = element_text(size = 10, color = "black"),
        text = element_text(size = 10)) +
  labs(y = "Normalised PFNA accumulation in resting assay [%]", x = "Normalised PFNA accumulation in growth assay [%]")
ggsave(file = "results/correlation_growth_resting/assay3_OD_corrected_correlation_outlier_removed.pdf",
       width = 10, height = 10, units = "cm")

rm(list = ls())
```



Gaussion micture models:
```{r eval = FALSE}
library(mclust)

# read in data
data <- read.table(file = "results/assay3_heatmap.txt", 
                   sep = "\t", dec = ".", header = TRUE)

# The Mclust function is used to fit Gaussian mixture models (G=2 components, allowing variances to be unequal (model="V")).
fit = Mclust(data$percent.accumulation, G=2, model="V")
summary(fit)

# plot the resulting density plot
plot(fit, what="density", main="", xlab="Percent accumulation [%}")

# filter out data, classification and uncertainty
fit_data <- data.frame(percent.accumulation = fit$data, 
                       classification = fit$classification, 
                       uncertainty = fit$uncertainty)

# merge data and fit_data
new_data <- merge(data, fit_data, by = "percent.accumulation")

new_data <- new_data %>%
  select(experiment, batch, strain.id, strain, percent.accumulation, classification, uncertainty) %>%
  arrange(strain)

write.table(new_data,
            file = "results/assay3_mixture_model.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

strains <- read.table(file = "data/strains.txt", 
                   sep = "\t", dec = ".", header = TRUE)
new_data_strains <- merge(new_data, strains, by = c("strain", "strain.id"), all = TRUE)

count(new_data_strains, classification)

rm(list = ls())
```



Species level abundance, prevalence, accumulation analysis:
```{r eval = FALSE}
data <- read.table(file = "results/abundance_prevalence/assay3_abundance_prevalence.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data$classification <- as.factor(data$classification)
levels(data$classification)
levels(data$classification) <- c("low-acccumulating species","high-accumulating species")


# distribution of abundance by class
ggplot(data, aes(x = abundance, color = classification)) +
  geom_density() +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 10, color = "black"),
        axis.title.y = element_text(size = 10, color = "black"),
        text = element_text(size = 10),
        legend.position = c(0.7, 0.8),
        legend.key.size = unit(0.6,"line"),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10)) +
  labs(x = "Abundance [%]", y = "Density", color = "Classification")
ggsave(file = "results/abundance_prevalence/assay3_distribution_abundance.png",
       width = 10, height = 10, units = "cm")

# distribution of prevalence by class
ggplot(data, aes(x = prevalence, color = classification)) +
  geom_density() +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 10, color = "black"),
        axis.title.y = element_text(size = 10, color = "black"),
        text = element_text(size = 10),
        legend.position = c(0.5, 0.1),
        legend.key.size = unit(0.6,"line"),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10)) +
  labs(x = "Prevalence [%]", y = "Density", color = "Classification")
ggsave(file = "results/abundance_prevalence/assay3_distribution_prevalence.png",
       width = 10, height = 10, units = "cm")


# correlation between prevalence and PFNA accumulation
cor.test(data$prevalence, data$percent.accumulation, method = "pearson")
# r = 0.17; p = 0.26; no significant correlation
cor.test(data$prevalence, data$percent.accumulation, method = "spearman")
# rho = 0.15; p = 0.29; no significant correlation

ggplot(data, aes(x = prevalence, y = percent.accumulation)) +
  geom_point() +
  geom_smooth(method = "lm", se = T) +
  stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")), size = 3, label.x = 1, label.y = 60) +
  annotate(geom="text", label="r = 0.17; p = 0.26", x=20, y=55, size = 3) +
  annotate(geom="text", label="rho = 0.15; p = 0.29", x=20, y=50, size = 3) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 10, color = "black"),
        axis.title.y = element_text(size = 10, color = "black"),
        text = element_text(size = 10)) +
  labs(x = "Prevalence [%]", y = "PFNA accumulation [%]")
ggsave(file = "results/abundance_prevalence/assay3_cor_accumulation_prevalence.pdf",
       width = 10, height = 10, units = "cm")


# correlation between abundance and PFNA accumulation
cor.test(data$abundance, data$percent.accumulation, method = "pearson")
# r = 0.2; p = 0.17; no significant correlation
cor.test(data$abundance, data$percent.accumulation, method = "spearman")
# rho = 0.21; p = 0.16; no significant correlation

ggplot(data, aes(x = abundance, y = percent.accumulation)) +
  geom_point() +
  geom_smooth(method = "lm", se = T) +
  stat_regline_equation(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~~")), size = 3, label.x = log10(0.2), label.y = 7) +
  annotate(geom="text", label="r = 0.17; p = 0.26", x=0.7, y=4, size = 3) +
  annotate(geom="text", label="rho = 0.15; p = 0.29", x=0.7, y=1, size = 3) +
  scale_x_log10() +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        axis.title.x = element_text(size = 10, color = "black"),
        axis.title.y = element_text(size = 10, color = "black"),
        text = element_text(size = 10)) +
  labs(x = "Abundance [%]", y = "PFNA accumulation [%]")
ggsave(file = "results/abundance_prevalence/assay3_cor_accumulation_abundance.pdf",
       width = 10, height = 10, units = "cm")

rm(list = ls())
```




Plot results summary:
```{r eval = FALSE}
strains <- read.table(file = "data/strains.txt", 
                   sep = "\t", dec = ".", header = TRUE)

mm <- read.table(file = "results/assay3_mixture_model.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data <- merge(mm, strains, by = c("strain", "strain.id"), all = TRUE)
data <- data %>%
  select(strain.id, strain, phylum, gram.stain, percent.accumulation, classification) %>%
  mutate_at(c("classification"), as.factor) %>%
  mutate(classification = recode_factor(classification, "1" = "low-accumulating", "2" = "high-accumulating"))

write.table(data,
            file = "results/assay3_results_summary.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

# make phylum pie chart
phylum <- data %>%
  count(phylum)
png(file = "results/assay3_phylum_piechart.pdf")
pie(x = phylum$n, labels = phylum$phylum, col = c("#a8a8a7", "#ee3f36", "#ee8300", "#8bc53f", "#f9ec31", "#ee00ee", "#00afee"), angle = 45)
dev.off()
ggplot(phylum, aes(x = "", y = n, fill = phylum)) +
  geom_bar(stat = "identity") +
  coord_polar("y", start = 0, direction = -1) +
  scale_fill_manual(values = c("#a8a8a7", "#ee3f36", "#ee8300", "#8bc53f", "#f9ec31", "#ee00ee", "#00afee")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),)
ggsave(file = "results/assay3_piechart_phylum.pdf",
       width = 10, height = 10, units = "cm")

# make phylum stacked bar chart
phylum <- data %>%
  count(phylum)
ggplot(phylum, aes(x = "", y = n, fill = phylum)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("#ee3f36", "#ee8300", "#8bc53f", "#f9ec31", "#ee00ee", "#00afee", "#a8a8a7")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),)
ggsave(file = "results/assay3_barplot_phylum.png",
       width = 10, height = 10, units = "cm")

# phylum heatmap
ggplot(data, aes(x = 1, y=fct_rev(strain))) +
  geom_tile(aes(fill = phylum), color = "black") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black", face = "italic"),
        text = element_text(size = 10)) +
  scale_fill_manual(values = c("#ee3f36", "#ee8300", "#8bc53f", "#f9ec31", "#ee00ee", "#00afee", "#a8a8a7"))
ggsave(file = "results/assay3_heatmap_phylum.png",
       width = 12, height = 50, units = "cm")

# gram-stain heatmap
ggplot(data, aes(x = 1, y=fct_rev(strain))) +
  geom_tile(aes(fill = gram.stain), color = "black") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black", face = "italic"),
        text = element_text(size = 10))
ggsave(file = "results/assay3_heatmap_gramstain.png",
       width = 12, height = 50, units = "cm")

# classification heatmap
ggplot(data, aes(x = 1, y=fct_rev(strain))) +
  geom_tile(aes(fill = classification), color = "black") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black", face = "italic"),
        text = element_text(size = 10)) +
  scale_fill_manual(values = c("#c9c9c9", "#424242"))
ggsave(file = "results/assay3_heatmap_class.pdf",
       width = 12, height = 50, units = "cm")
```
