---
title: "PFAS pellet recovery"
author: "Anna Lindell"
date: "Oktober 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(tidyverse)
library(readr)
library(ggh4x)
```





# Quality control

Internal standard:
```{r eval = FALSE}
data <- read.table(file = "data/fu11_data_raw.txt", 
                   sep = "\t", dec = ".", header = TRUE)

is <- data %>% filter(group != "Blank")

ggplot(is, aes(x = file, y = is.resp)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_blank()) +
  facet_grid(cols = vars(media)) +
  labs(y = "Internal Standard Response", x = "Data File") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/fu11_is.png",
       width = 15, height = 10, units = "cm")

print.data.frame(is %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), is.resp))

rm(list = ls())
```

Internal standard Caffeine is different for mGAM and PBS and pellet and supernatant --> matrix effects
Coefficient of variation: 45 %




Quality control for response:
```{r eval = FALSE}
qc <- read.table(file = "data/fu11_qc.txt", 
                   sep = "\t", dec = ".", header = TRUE)

ggplot(qc, aes(x = number, y = compound.resp)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_blank()) +
  facet_grid(cols = vars(compound), rows = vars(media), scales = "free_x") +
  labs(y = "Response", x = "Data File") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/fu11_qc_resp.png",
       width = 15, height = 10, units = "cm")

ggplot(qc, aes(x = number, y = ratio)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_blank()) +
  facet_grid(cols = vars(compound), rows = vars(media), scales = "free_x") +
  labs(y = "Ratio", x = "Data File") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/fu11_qc_ratio.png",
       width = 15, height = 10, units = "cm")

print.data.frame(qc %>% group_by(compound, media) %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), compound.resp))
print.data.frame(qc %>% group_by(compound, media) %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), ratio))

rm(list = ls())
```

| Compound | Matrix | cv Response | cv Ratio |
| -------- | ------ | ----------- | -------- |
| PFHpA    | PBS    | 2 %         | 3.3 %    |
| PFHpA    | mGAM   | 2.2 %       | 3.6 %    |
| PFOA     | PBS    | 2.7 %       | 3.9 %    |
| PFOA     | mGAM   | 1.6 %       | 3 %      |
| PFNA     | PBS    | 3.2 %       | 4.1 %    |
| PFNA     | mGAM   | 2.2 %       | 3.7 %    |
| PFDA     | PBS    | 3.6 %       | 4.8 %    |
| PFDA     | mGAM   | 10.6 %      | 10.4 %   |

Response used for concentration calculations.





# Median Reduction:

```{r eval = FALSE}
data <- read.table(file = "data/fu11_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)

cc_median <- data %>%
  filter(group == "Compound control") %>%
  group_by(media, compound) %>%
  summarise(median = median(conc)) 

n <- data %>%
  group_by(media, compound) %>%
  count()
n <- n$n

cc_median <- cc_median$median
cc_median <- rep(cc_median,  n)

median <- data %>%
  arrange(experiment, media, compound, strain, group, replicate) %>%
  mutate(percent = conc / cc_median * 100)

write.table(median,
            file = "data/fu11_medianreduction.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```




# Plot samples

cc, p, sn:
```{r eval = FALSE}
data <- read.table(file = "data/fu11_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data$strain <- as.factor(data$strain)
levels(data$strain)
data$strain <- factor(data$strain, levels=c("Compound control", "B. uniformis", "E. coli BW25113", "E. coli BW25113 delta-tolC"))
data$compound <- as.factor(data$compound)
levels(data$compound)
data$compound <- factor(data$compound, levels=c("PFHpA", "PFOA", "PFNA", "PFDA"))

pbs <- data %>%
  filter(media == "PBS")
mgam <- data %>%
  filter(media == "mGAM")

ggplot(pbs, aes(x = group, y = percent)) +
  geom_jitter(width = 0.1, size = 3, shape = 1) +
  facet_grid2(rows = vars(compound), cols = vars(strain), scales = "free_x", labeller = label_wrap_gen(width = 13, multi_line = TRUE)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 0.5, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        strip.background = element_blank()) +
  labs(y = "PFAS recovery [%]") +
  ylim(0,NA)
ggsave(file = "results/fu11_pbs_percent.pdf",
           width = 10, height = 15, units = "cm")

ggplot(mgam, aes(x = group, y = percent)) +
  geom_jitter(width = 0.1, size = 3, shape = 1) +
  facet_grid2(rows = vars(compound), cols = vars(strain), scales = "free_x", labeller = label_wrap_gen(width = 13, multi_line = TRUE)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 0.5, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        strip.background = element_blank()) +
  labs(y = "PFAS recovery [%]") +
  ylim(0,NA)
ggsave(file = "results/fu11_mgam_percent.pdf",
           width = 10, height = 15, units = "cm")

data$group <- as.factor(data$group)
levels(data$group)
data$group <- factor(data$group, levels=c("Compound control", "Supernatant", "Pellet"))

data.sum <- data %>%
  group_by(media, compound, strain, group) %>%
  summarise(se = sd(percent)/length(percent),
            percent = median(percent))
error_bars <- data.sum %>%
  arrange(media, compound, strain, desc(group)) %>%
  group_by(media, compound, strain) %>%
  mutate(percent_new = cumsum(percent))

ggplot(data.sum, aes(x = strain, y = percent, fill = group)) +
  geom_bar(stat = "identity") +
  geom_errorbar(data = error_bars,
                aes(x = strain, ymax = percent_new + se, ymin = percent_new - se), 
                width = 0.2) +
  facet_grid(rows = vars(media), cols = vars(compound), scales = "free_x") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 0.5, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        strip.background = element_blank()) +
  labs(y = "PFAS recovery [%]") +
  ylim(0,NA)
ggsave(file = "results/fu11_mgam_percent_barplot.pdf",
           width = 15, height = 11, units = "cm")

rm(list = ls())
```




# Statistical analyis

T-test to test if supernatant significantly reduced:
```{r eval = FALSE}
data <- read.table(file = "data/fu11_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

samples <- data %>% filter(group == c("Supernatant"))
controls <- data %>% filter(group == "Compound control")

med <- unique(samples$media)
comp <- unique(samples$compound)
strains <- unique(samples$strain)

n <- 1
p_value <- data.frame(media = character(),
                      compound = character(),
                      strain = character(),
                      p.value = numeric())
for (m in seq_along(med)) {
  samples1 <- samples %>% filter(media == med[m])
  controls1 <- controls %>% filter(media == med[m])
  for (c in seq_along(comp)) {
    samples2 <- samples1 %>% filter(compound == comp[c])
    controls2 <- controls1 %>% filter(compound == comp[c])
    for (s in seq_along(strains)) {
      samples3 <- samples2 %>% filter(strain == strains[s])
      test <- t.test(x = samples3$conc, y = controls2$conc, alternative = "two.sided")
      p_value[n, 1:4] <- c(med[m], comp[c], strains[s], test$p.value)
      n <- n + 1
    }
  }
}

adj_data <- p_value %>%
  group_by(media, strain) %>%
  mutate(p_adj = p.adjust(p.value, method = "fdr")) %>%
  arrange(media, compound)

write.table(adj_data,
            file = "results/fu11_p_value_sn.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)
```


Results summary:
```{r eval = FALSE}
median <- data %>%
  group_by(experiment, media, compound, strain.id, strain, group, unit) %>%
  summarise(conc = median(conc),
            percent = median(percent))

write.table(median,
            file = "results/fu11_results_summary.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```




# Results analysis:

Check addition of pellet and supernatant concentration:
```{r eval = FALSE}
data <- read.table(file = "data/fu11_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

new.data <- data %>%
  filter(strain.id != "Compound control") %>%
  pivot_wider(id_cols = c("media", "compound", "strain.id", "strain", "replicate"), names_from = "group", values_from = "percent") %>%
  mutate(Sum = Pellet + Supernatant)

new.data.sum <- new.data %>%
  group_by(media, compound, strain.id, strain) %>%
  summarise(Pellet = median(Pellet),
                 Supernatant = median(Supernatant)) %>%
  mutate(Sum = Pellet + Supernatant)

write.table(new.data,
            file = "results/fu11_results_sum_calculation.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)
write.table(new.data.sum,
            file = "results/fu11_results_sum_calculation_summary.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```



Calculate concentration in pellet based on pellet weight:
```{r eval = FALSE}
data <- read.table(file = "data/fu11_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

mw <- data_frame(compound = c("PFHpA", "PFOA", "PFNA", "PFDA"),
                 molecular.weight = c(364.06, 414.07, 464.08, 514.08))

control <- data %>% 
  filter(group == "Compound control") %>%
  group_by(media, compound) %>%
  summarise(control.conc = median(conc))

pellet.weight <- read.table(file = "data/fu11_pellet_weight.txt", 
                   sep = "\t", dec = ".", header = TRUE)

pellet.pbs <- data %>%
  filter(group == "Pellet") %>%
  select(experiment, media, compound, strain.id, strain, replicate, group, conc, percent) %>%
  full_join(mw, by = c("compound")) %>%
  full_join(control, by = c("media", "compound")) %>%
  full_join(pellet.weight, by = c("media", "compound", "strain.id", "strain", "replicate")) %>%
  select(-c(tube.mg, tubepellet.mg)) %>%
  filter(media == "PBS") %>%
  mutate(pellet.uM.1 = conc * 0.0005 / pellet.mg * 1000000, # volume of culture was 0.5 ml
         foldchange.1 = pellet.uM.1 / control.conc,
         pellet.ugl.1 = pellet.uM.1 * molecular.weight,
         pellet.uM.2 = conc * 0.0005 / (3.75 * 1.3*10^9 / 2 * 1 * 1E-15), # volume of culture was 0.5 ml; 3.75 was OD; average of 1.3*10^9 cells per OD unit; 1um3 volume per cell
         foldchange.2 = pellet.uM.2 / control.conc,
         pellet.ugl.2 = pellet.uM.2 * molecular.weight)

pellet.mgam <- data %>%
  filter(group == "Pellet") %>%
  select(experiment, media, compound, strain.id, strain, replicate, group, conc, percent) %>%
  full_join(mw, by = c("compound")) %>%
  full_join(control, by = c("media", "compound")) %>%
  full_join(pellet.weight, by = c("media", "compound", "strain.id", "strain", "replicate")) %>%
  filter(media == "mGAM") %>%
  select(-c(tube.mg, tubepellet.mg)) %>%
  mutate(pellet.uM.1 = conc * 0.0005 / pellet.mg * 1000000, # volume of culture was 0.5 ml
         foldchange.1 = pellet.uM.1 / control.conc,
         pellet.ugl.1 = pellet.uM.1 * molecular.weight)

pellet <- bind_rows(pellet.pbs, pellet.mgam)

write.table(pellet,
            file = "results/fu11_results_pellet_concentration.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

pellet.sum <- pellet %>%
  group_by(experiment, media, compound, strain.id, strain, group) %>%
  summarise(conc = median(conc),
            percent = median(percent),
            molecular.weight = median(molecular.weight),
            control.conc = median(control.conc),
            pellet.mg = median(pellet.mg),
            pellet.uM.1 = median(pellet.uM.1),
            foldchange.1 = median(foldchange.1),
            pellet.ugl.1 = median(pellet.ugl.1),
            pellet.uM.2 = median(pellet.uM.2),
            foldchange.2 = median(foldchange.2),
            pellet.ugl.2 = median(pellet.ugl.2))

write.table(pellet.sum,
            file = "results/fu11_results_pellet_concentration_summary.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

# heatmap
pellet.sum$compound <- as.factor(pellet.sum$compound)
levels(pellet.sum$compound)
pellet.sum$compound <- factor(pellet.sum$compound, levels = c("PFHpA", "PFOA", "PFNA", "PFDA"))

pellet.sum$strain <- as.factor(pellet.sum$strain)
levels(pellet.sum$strain)
pellet.sum$strain <- factor(pellet.sum$strain, levels = c("E. coli BW25113 delta-tolC", "E. coli BW25113", "B. uniformis"))

ggplot(pellet.sum, aes(x = compound, y = strain)) +
  geom_tile(aes(fill = foldchange.1), color = "black") +
  geom_text(aes(label = round(foldchange.1, 0)), size = 3) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, color = "black"),
        axis.text.y = element_text(color = "black"),
        text = element_text(size = 10),
        strip.background = element_blank()) +
  scale_fill_gradient(name = "Foldchange", low = "white", high = "#546bab") +
  facet_wrap(~media)
ggsave("results/fu11_heatmap_foldchange_1.pdf",
       width = 11, height = 4, units = "cm")

make_triangles <- function(x, y, point = "up") {
  x <- as.integer(as.factor((x)))
  y <- as.integer(as.factor((y)))

  if (point == "up") {
    newx <- sapply(x, function(x) {
      c(x - 0.5, x - 0.5, x + 0.5)
    }, simplify = FALSE)
    newy <- sapply(y, function(y) {
      c(y - 0.5, y + 0.5, y + 0.5)
    }, simplify = FALSE)
  } else if (point == "down") {
    newx <- sapply(x, function(x) {
      c(x - 0.5, x + 0.5, x + 0.5)
    }, simplify = FALSE)
    newy <- sapply(y, function(y) {
      c(y - 0.5, y - 0.5, y + 0.5)
    }, simplify = FALSE)
  }
  data.frame(x = unlist(newx), y = unlist(newy))
}

newcoord_up <- make_triangles(pellet.sum$compound, pellet.sum$strain)
newcoord_down <- make_triangles(pellet.sum$compound, pellet.sum$strain, point = "down")
newcoord_down <- newcoord_down %>% select(xdown = x, ydown = y)  # rename columns

# you need to repeat each row of your previous data frame 3 times
repdata <- map_df(1:nrow(pellet.sum), function(i) pellet.sum[rep(i, 3), ])
newdata <- bind_cols(repdata, newcoord_up, newcoord_down)

ggplot(newdata) +
  geom_polygon(aes(x = x, y = y, fill = foldchange.1, group = interaction(compound, strain)), color = "black", linewidth = 0.1) +
  geom_polygon(aes(x = xdown, y = ydown, fill = foldchange.2, group = interaction(compound, strain)), color = "black", linewidth = 0.1) +
  scale_x_continuous(breaks = seq_along(unique(pellet.sum$compound)), 
                     labels = unique(levels(pellet.sum$compound))) +
  scale_y_continuous(breaks = seq_along(unique(pellet.sum$strain)),
                     labels = unique(levels(pellet.sum$strain)))+
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, color = "black"),
        axis.text.y = element_text(color = "black"),
        text = element_text(size = 10),
        strip.background = element_blank()) +
  scale_fill_gradient(name = "Foldchange", low = "white", high = "#546bab") +
  facet_wrap(~media)
ggsave("results/fu11_heatmap_foldchange_2.pdf",
       width = 11, height = 4, units = "cm")

```





