---
title: "PFAS accumulation in ng/l range"
author: "Anna Lindell"
date: "December 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(tidyverse)
library(readr)
library(ggh4x)
```


# Internal Standard

```{r eval = FALSE}
data <- read.table(file = "data/fu19_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)
```


Internal standard:
```{r eval = FALSE}
is <- data %>% add_column(number = 1:nrow(data)) %>%
  filter(group != "Blank")

ggplot(is, aes(x = number, y = is.resp)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_blank()) +
  labs(y = "Internal Standard Response", x = "Data File") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/fu19_is.png",
       width = 10, height = 7, units = "cm")

print.data.frame(is %>% 
  summarise_each(funs(median, mean, sd, cv=sd/mean), is.resp))

rm(list = ls())
```

 % cv for the Internal standard M9PFNA: 79 %



# Add concentration coulmn:

The internal standard was added at a concentration of 100 ng/l, therefore the ratio gives us the concentration in the samples.

```{r eval = FALSE}
data <- read.table(file = "data/fu19_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data <- data %>%
  mutate(unit = "ng/l",
         conc = ratio *100)


# data from imperial
data.imp <- read.table(file = "data/fu19_data_imperial.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data.imp <- data.imp %>%
  mutate(unit = "ng/l",
         conc = ratio *100)
```


# Median Reduction:

```{r eval = FALSE}
data <- data %>%
  filter(group != "Blank", group != "Blank solvent" ) %>%
  group_by(experiment, strain.id, strain, group, replicate, compound, is, unit) %>%
  summarize(compound.resp = mean(compound.resp),
            is.resp = mean(is.resp),
            ratio = mean(ratio),
            conc = mean(conc))


cc_median <- data %>%
  filter(group == "Compound control") %>%
  group_by(compound) %>%
  summarise(median = median(ratio)) 

cc_median <- cc_median$median

median <- data %>%
  arrange(experiment, strain, group, replicate) %>%
  mutate(percent = ratio / cc_median * 100)

write.table(median,
            file = "data/fu19_medianreduction.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)


# data from Imperial
data.imp <- data.imp %>%
  filter(group != "Blank", group != "Standard" ) %>%
  group_by(experiment, strain.id, strain, group, replicate, compound, is, unit) %>%
  summarize(compound.resp = mean(compound.resp),
            is.resp = mean(is.resp),
            ratio = mean(ratio),
            conc = mean(conc))


cc_median.imp <- data.imp %>%
  filter(group == "Compound control") %>%
  group_by(compound) %>%
  summarise(median = median(ratio)) 

cc_median.imp <- cc_median.imp$median

median.imp <- data.imp %>%
  arrange(experiment, strain, group, replicate) %>%
  mutate(percent = ratio / cc_median.imp * 100)

write.table(median.imp,
            file = "data/fu19_medianreduction_imperial.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```

# calculate concentration in pellets and plot samples

```{r eval = FALSE}
data <- read.table(file = "data/fu19_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

pw <- read.table(file = "data/fu19_pellet_weight.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data <- data %>%
  select(experiment, strain.id, strain, group, replicate, compound, is, unit, conc, percent) %>%
  full_join(pw, by = c("experiment", "strain.id", "strain", "group", "replicate"))

sn <- data %>%
  filter(group %in% c("Supernatant", "Compound control", "Extractant blank", "PBS blank"))

pellet <- data %>%
  filter(group %in% c("Pellet", "Pellet blank")) %>%
  mutate(conc = conc * 0.5 / sample.weight.g * 1000)

data.new <- rbind(sn, pellet)


data.new$group <- as.factor(data.new$group)
levels(data.new$group)
data.new$group <- factor(data.new$group, levels = c("Extractant blank", "Pellet blank", "Pellet", "PBS blank", "Supernatant", "Compound control"))

ggplot(data.new, aes(x = group, y = percent)) +
  geom_jitter(width = 0.1, size = 2, shape = 1) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 1, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        strip.background = element_blank()) +
  labs(y = "PFNA recovery [%]")
ggsave(file = "results/fu19_pfna_percent.png",
       width = 6, height = 8, units = "cm")

ggplot(data.new, aes(x = group, y = conc)) +
  geom_jitter(width = 0.1, size = 2, shape = 1) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 1, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        strip.background = element_blank()) +
  scale_y_log10() +
  labs(y = "PFNA recovery [ng/l]")
ggsave(file = "results/fu19_pfna_conc.pdf",
       width = 6, height = 8, units = "cm")

sam <- data.new %>% filter(group %in% c("Pellet blank", "Pellet", "Supernatant", "Compound control"))
ggplot(sam, aes(x = group, y = conc, shape = group)) +
  geom_jitter(width = 0.1, size = 3) +
  scale_shape_manual(values = c(2,2,1,3)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 1, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        legend.position = "none",
        strip.background = element_blank()) +
  labs(y = "PFNA recovery [ng/l]") +
  scale_y_log10()
ggsave(file = "results/fu19_pfna_conc_subset.pdf",
       width = 6, height = 8, units = "cm")

ggplot(sam, aes(x = group, y = percent, shape = group)) +
  geom_jitter(width = 0.1, size = 3) +
  scale_shape_manual(values = c(2,2,1,3)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 1, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        legend.position = "none",
        strip.background = element_blank()) +
  ylim(0,NA) +
  labs(y = "PFNA recovery [%]")
ggsave(file = "results/fu19_pfna_percent_subset.png",
       width = 6, height = 8, units = "cm")

write.table(data.new,
            file = "results/fu19_results.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

median <- data.new %>%
  group_by(strain.id, strain, group, compound, is, unit) %>%
  summarize(sample.weight.g = mean(sample.weight.g),
            conc = median(conc),
            percent = median(percent))


write.table(median,
            file = "results/fu19_results_summary.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```


Plot samles from Imperial:
```{r eval = FALSE}
data <- read.table(file = "data/fu19_medianreduction_imperial.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data$group <- as.factor(data$group)
levels(data$group)
data$group <- factor(data$group, levels = c("Extractant blank", "Pellet blank", "Pellet", "PBS blank", "Supernatant", "Compound control"))

ggplot(data, aes(x = group, y = ratio)) +
  geom_jitter(width = 0.1, size = 2, shape = 1) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 1, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        strip.background = element_blank()) +
  labs(y = "Response Ratio (PFNA / M9PFNA)") 
ggsave(file = "results/fu19_pfna_ratio_imperial.png",
       width = 6, height = 8, units = "cm")

pw <- read.table(file = "data/fu19_pellet_weight.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data <- data %>%
  select(experiment, strain.id, strain, group, replicate, compound, is, unit, conc, percent) %>%
  full_join(pw, by = c("strain.id", "strain", "group", "replicate"))

sn <- data %>%
  filter(group %in% c("Supernatant", "Compound control", "Extractant blank", "PBS blank"))

pellet <- data %>%
  filter(group %in% c("Pellet", "Pellet blank")) %>%
  mutate(conc = conc * 0.5 / sample.weight.g * 1000)

data.new <- rbind(sn, pellet)

data.new$group <- as.factor(data.new$group)
levels(data.new$group)
data.new$group <- factor(data.new$group, levels = c("Extractant blank", "Pellet blank", "Pellet", "PBS blank", "Supernatant", "Compound control"))

data.new$replicate <- as.factor(data.new$replicate)

ggplot(data.new, aes(x = group, y = percent)) +
  geom_jitter(width = 0.1, size = 2, shape = 1) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 1, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        strip.background = element_blank()) +
  labs(y = "PFNA recovery [%]") 
ggsave(file = "results/fu19_pfna_percent_imperial.png",
       width = 6, height = 8, units = "cm")

ggplot(data.new, aes(x = group, y = conc)) +
  geom_jitter(width = 0.1, size = 2, shape = 1) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 1, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        strip.background = element_blank()) +
  labs(y = "PFNA recovery [ng/l]") +
  scale_y_log10()
ggsave(file = "results/fu19_pfna_conc_imperial.pdf",
       width = 6, height = 8, units = "cm")


sam <- data.new %>% filter(group %in% c("Pellet blank", "Pellet", "Supernatant", "Compound control"))
ggplot(sam, aes(x = group, y = conc, shape = group)) +
  geom_jitter(width = 0.1, size = 3) +
  scale_shape_manual(values = c(2,2,1,3)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 1, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        legend.position = "none",
        strip.background = element_blank()) +
  labs(y = "PFNA recovery [ng/l]") +
  scale_y_log10()
ggsave(file = "results/fu19_pfna_conc_subset_imperial.png",
       width = 6, height = 8, units = "cm")

ggplot(sam, aes(x = group, y = percent, shape = group)) +
  geom_jitter(width = 0.1, size = 3) +
  scale_shape_manual(values = c(2,2,1,3)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 1, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        legend.position = "none",
        strip.background = element_blank()) +
  ylim(0,NA) +
  labs(y = "PFNA recovery [%]")
ggsave(file = "results/fu19_pfna_percent_subset_imperial.pdf",
       width = 6, height = 8, units = "cm")

write.table(data.new,
            file = "results/fu19_results_imperial.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

median <- data.new %>%
  group_by(strain.id, strain, group, compound, is, unit) %>%
  summarize(sample.weight.g = mean(sample.weight.g),
            conc = median(conc),
            percent = median(percent))

write.table(median,
            file = "results/fu19_results_summary_imperial.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```


# Statistical analyis

T-test to test if supernatant significantly reduced and pellet significantly increased:
```{r eval = FALSE}
data <- read.table(file = "data/fu19_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

pelletcontrol <- data %>% filter(group == c("Pellet blank"))
pelletsamples <- data %>% filter(group == c("Pellet"))
compoundcontrol <- data %>% filter(group == c("Compound control"))
snsamples <- data %>% filter(group == c("Supernatant"))

t.test(x = pelletsamples$conc, y = pelletcontrol$conc, alternative = "two.sided")
t.test(x = snsamples$conc, y = compoundcontrol$conc, alternative = "two.sided")

t.test(x = pelletsamples$percent, y = pelletcontrol$percent, alternative = "two.sided")
t.test(x = snsamples$percent, y = compoundcontrol$percent, alternative = "two.sided")

```

Concentration & percent:
Pellet concentration significantly increased: p = 0.002583
Supernatant concentration significantly decreased compared to compound control: p = 0.02016



T-test to test if supernatant significantly reduced and pellet significantly increased:
```{r eval = FALSE}
data <- read.table(file = "data/fu19_medianreduction_imperial.txt", 
                   sep = "\t", dec = ".", header = TRUE)

pelletcontrol <- data %>% filter(group == c("Pellet blank"))
pelletsamples <- data %>% filter(group == c("Pellet"))
compoundcontrol <- data %>% filter(group == c("Compound control"))
snsamples <- data %>% filter(group == c("Supernatant"))

t.test(x = pelletsamples$conc, y = pelletcontrol$conc, alternative = "two.sided")
t.test(x = snsamples$conc, y = compoundcontrol$conc, alternative = "two.sided")

t.test(x = pelletsamples$percent, y = pelletcontrol$percent, alternative = "two.sided")
t.test(x = snsamples$percent, y = compoundcontrol$percent, alternative = "two.sided")

```
