---
title: "PFNA-B. uniformis 7d resting assay"
author: "Anna Lindell"
date: "20/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(tidyverse)
library(readr)
```




# Quality control

Read in data:
```{r eval = FALSE}
data <- read.table(file = "data/fu3_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)
```


Internal standard:
```{r eval = FALSE}
is <- data %>% filter(group != "Blank")
is <- is[!duplicated(is$file), ]

ggplot(is, aes(x = file, y = is.resp)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_rect(fill="white")) +
  labs(y = "Internal Standard Response", x = "Data File") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/fu3_is.png",
       width = 10, height = 7, units = "cm")

print.data.frame(is %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), is.resp))
```

Ibuprofen: Coefficient of variation 6.5 %
-> raw response used for further calculations


Quality control:
```{r eval = FALSE}
qc <- data %>% filter(group == "QC")

ggplot(qc, aes(x = file, y = compound.resp)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_rect(fill="white")) +
  labs(y = "Internal Standard Response", x = "Data File") +
  facet_wrap(~compound, scales = "free_x") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/fu3_qc.png",
       width = 7, height = 5, units = "cm")

print.data.frame(qc %>% group_by(compound) %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), compound.resp))

rm(list = ls())
```

QC: Coefficient of variation 3.4 %
-> no significant signal drift





# Median Reduction:

```{r eval = FALSE}
data <- read.table(file = "data/fu3_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data <- data %>%
  filter(group %in% c("Compound control","Supernatant","Pellet","Whole culture"))

cc_median <- data %>%
  filter(group == "Compound control") %>%
  group_by(timepoint.d) %>%
  summarise(median = median(conc))

cc_median <- cc_median$median
cc_median <- rep(cc_median,  each = 12)

median <- data %>%
  arrange(timepoint.d, strain, group, replicate) %>%
  mutate(percent = conc / cc_median * 100)

write.table(median,
            file = "data/fu3_medianreduction.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```




# Plot samples

```{r eval = FALSE}
data <- read.table(file = "data/fu3_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data$group <- as.factor(data$group)
levels(data$group)
data$group <- factor(data$group, levels = c("Compound control", "Supernatant", "Pellet", "Whole culture"))

ggplot(data, aes(x = group, y = conc)) +
  geom_jitter(width = 0.1, size = 3, shape = 1) +
  facet_grid(~timepoint.d, scales = "free_x") +
  theme_bw() +
  ylim(0, NA) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_text(face = "italic"),
        axis.text.x = element_text(color = "black", angle = 90, , vjust = 0.5, hjust = 1),
        axis.text.y = element_text(color = "black"),
        strip.background = element_rect(colour="white", fill="white")) +
  labs(y = "PFNA recovery [uM]", x = "B. unifomris")
ggsave(file = "results/fu3_concentration.png",
           width = 15, height = 7, units = "cm")

ggplot(data, aes(x = group, y = percent)) +
  geom_jitter(width = 0.1, size = 3, shape = 1) +
  facet_grid(~timepoint.d, scales = "free_x") +
  theme_bw() +
  ylim(0, 135) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_text(face = "italic"),
        axis.text.x = element_text(color = "black", angle = 90, , vjust = 0.5, hjust = 1),
        axis.text.y = element_text(color = "black"),
        strip.background = element_rect(colour="white", fill="white")) +
  labs(y = "PFNA recovery [%]", x = "B. unifomris")
ggsave(file = "results/fu3_percent.png",
           width = 15, height = 7, units = "cm")

median1 <- data %>%
  group_by(compound, timepoint.d, group) %>%
  summarise(conc = median(conc),
            percent = median(percent))

ggplot(median1, aes(x = timepoint.d, y = conc, shape = group)) +
  geom_line(aes(linetype = group), size = 0.5, color = "#808080") +
  geom_point(data = data, size = 3) +
  scale_shape_manual(values=c(3, 1, 2, 0)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_text(color = "black"),
        axis.text.y = element_text(color = "black")) +
  labs(y = "PFNA recovery [uM]", x = "Time [days]") +
  ylim(0, NA)
ggsave(file = "results/fu3_pfna-concentration-over-time.png",
           width = 14, height = 10, units = "cm")

ggplot(median1, aes(x = timepoint.d, y = percent, shape = group)) +
  geom_line(aes(linetype = group), size = 0.5, color = "#808080") +
  geom_point(data = data, size = 3) +
  scale_shape_manual(values=c(3, 1, 2, 0)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black")) +
  labs(y = "PFNA recovery [%]", x = "Time [days]") +
  ylim(0, NA)
ggsave(file = "results/fu3_pfna-percentage-over-time.png",
           width = 11, height = 7, units = "cm")


rm(list = ls())
```



# Statistical analyis

T-test to test if supernatant significantly reduced:
```{r eval = FALSE}
data <- read.table(file = "data/fu3_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

samples <- data %>% filter(group %in% c("Supernatant"))
controls <- data %>% filter(group == "Compound control")

time <- unique(data$timepoint.d)

n <- 1
p_value <- data.frame(time = character(),
                      p.value = numeric())
for (t in seq_along(time)) {
  samples1 <- samples %>% filter(timepoint.d == time[t])
  controls1 <- controls %>% filter(timepoint.d == time[t])
  test <- t.test(x = samples1$conc, y = controls1$conc, alternative = "two.sided")
  p_value[n, 1:2] <- c(time[t], test$p.value)
  n <- n + 1
}

write.table(p_value,
            file = "results/fu3_p_value_sn.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```



T-test to test if pellet significantly increased:
```{r eval = FALSE}
data <- read.table(file = "data/fu3_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

samples <- data %>% filter(group %in% c("Pellet"))

time <- unique(data$timepoint.d)

n <- 1
p_value <- data.frame(time = character(),
                      p.value = numeric())
for (t in seq_along(time)) {
  samples1 <- samples %>% filter(timepoint.d == time[t])
  test <- t.test(x = samples1$conc, mu = 0, alternative = "two.sided")
  p_value[n, 1:2] <- c(time[t], test$p.value)
  n <- n + 1
}

write.table(p_value,
            file = "results/fu3_p_value_pellet.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```



Results summary:
```{r eval = FALSE}
median <- data %>%
  group_by(experiment, timepoint.d, strain.id, strain, group, compound) %>%
  summarise(conc = median(conc),
            percent = median(percent))

write.table(median,
            file = "results/fu3_results_summary.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)
```


PFNA is significantly reduced to at all tested time points. PFNA was not released by the cell after a while an accumulated amount did not change over time.




