---
title: "NT5002 & BWdC - PFNA accumulation at different ODs"
author: "Anna Lindell"
date: "8/08/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(tidyverse)
library(readr)
library(ggpubr)
library(ggplot2)
library(ggpp)
library(ggpmisc)
```



# Quality control

Read in data:
```{r eval = FALSE}
data <- read.table(file = "data/fu6_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)
```


Internal standard:
```{r eval = FALSE}
is <- data %>% filter(group != "Blank")

ggplot(is, aes(x = file, y = is.resp)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_rect(fill="white")) +
  labs(y = "Internal Standard Response", x = "Data File") +
  facet_wrap(~experiment) +
  ylim(0, NA)
ggsave(file = "qualitycontrol/fu6_is.png",
       width = 12, height = 7, units = "cm")

print.data.frame(is %>% group_by(experiment) %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), is.resp))
```

Caffeine: Coefficient of variation 
BWdC: 10.6 %
NT5002: 8.5 %
-> raw response used for further calculations



Quality control:
```{r eval = FALSE}
qc <- data %>% filter(group == "QC")

ggplot(qc, aes(x = file, y = compound.resp)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_rect(fill="white")) +
  labs(y = "Response", x = "Data File") +
  facet_wrap(~experiment, scales = "free_x") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/fu6_qc.png",
       width = 7, height = 5, units = "cm")

print.data.frame(qc %>% group_by(experiment) %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), compound.resp))

rm(list = ls())
```

QC: Coefficient of variation 
BWdC: 1.6 %
NT5002: 14.4
-> no signal drift





# Median Reduction:

```{r eval = FALSE}
data <- read.table(file = "data/fu6_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data <- data %>%
  filter(group %in% c("Compound control","Supernatant","Pellet","Whole culture")) 


cc_median <- data %>%
  filter(group == "Compound control") %>%
  group_by(experiment) %>%
  summarise(median = median(conc))

cc_median <- cc_median$median
cc_median <- rep(cc_median,  c(96, 136))

median <- data %>%
  arrange(experiment, strain, assay, od, group, replicate) %>%
  mutate(percent = conc / cc_median * 100)

write.table(median,
            file = "data/fu6_medianreduction.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```




# Plot samples

Plot samples by assay:
```{r eval = FALSE}
data <- read.table(file = "data/fu6_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data$od <- as.factor(data$od)
levels(data$od)
data$od <- factor(data$od, levels=c("OD8", "OD7", "OD6", "OD5", "OD4", "OD3", "OD2", "OD1", ""))


live <- data %>% filter(assay == "live" | group == "Compound control")
lysed <- data %>% filter(assay == "lysed" | group == "Compound control")

library(ggh4x)
ggplot(live, aes(x = group, y = percent)) +
  geom_jitter(width = 0.1, size = 3, shape = 1) +
  facet_grid2(rows = vars(experiment), cols = vars(od), scales = "free_x", render_empty = FALSE) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 0.5, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        strip.background = element_rect(colour="white", fill="white")) +
  labs(y = "PFNA recovery [%]") +
  ylim(0,NA)
ggsave(file = "results/fu6_live.png",
           width = 20, height = 10, units = "cm")

ggplot(lysed, aes(x = group, y = percent)) +
  geom_jitter(width = 0.1, size = 3, shape = 1) +
  facet_grid2(rows = vars(experiment), cols = vars(od), scales = "free_x", render_empty = FALSE) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 0.5, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        strip.background = element_rect(colour="white", fill="white")) +
  labs(y = "PFNA recovery [%]") +
  ylim(0,NA)
ggsave(file = "results/fu6_lysed.png",
           width = 20, height = 10, units = "cm")

buni <- live %>% filter(strain.id != "BWdC")
ggplot(buni, aes(x = percent, y = od, shape = group)) +
  geom_jitter(height = 0.1, size = 3) +
  scale_shape_manual(values = c(3,1,0)) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 0.5, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        strip.background = element_rect(colour="white", fill="white")) +
  labs(x = "PFNA recovery [%]") +
  xlim(0,NA)
ggsave(file = "results/fu6_nt5002_live.pdf",
           width = 11, height = 8, units = "cm")
```


Plot sn depletion:
```{r eval = FALSE}
sn <- data %>% filter(group == "Supernatant")
sn$od <- as.factor(sn$od)
levels(sn$od)
sn$od <- factor(sn$od, levels=c("OD8", "OD7", "OD6", "OD5", "OD4", "OD3", "OD2", "OD1"))

ggplot(sn, aes(x = percent, y = od, shape = assay)) +
  geom_jitter(height = 0.1, size = 3) +
  scale_shape_manual(values=c(1, 4)) +
  theme_bw() +
  xlim(0,100) +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        strip.background = element_blank()) +
  facet_grid(cols = vars(strain)) +
  labs(x = "Supernatant Recovery [%]")
ggsave(file = "results/fu6_percent_sn_livelysed.png",
           width = 12, height = 7, units = "cm")


snlive <- sn %>% filter(assay == "live")

ggplot(snlive, aes(x = percent, y = od)) +
  geom_point(size = 3, shape = 1) +
  theme_bw() +
  xlim(0,100) +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        strip.background = element_blank()) +
  facet_grid(cols = vars(strain)) +
  labs(x = "Supernatant Recovery [%]")
ggsave(file = "results/fu6_percent_sn_live.pdf",
           width = 12, height = 7, units = "cm")

rm(list = ls())
```



# Statistical analyis

T-test to test if supernatant significantly reduced:
```{r eval = FALSE}
data <- read.table(file = "data/fu6_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

samples <- data %>% filter(group == c("Supernatant"))
controls <- data %>% filter(group == "Compound control")

exp <- unique(samples$experiment)
as <- unique(samples$assay)
ods <- unique(samples$od)

n <- 1
p_value <- data.frame(experiment = character(),
                      assay = character(),
                      od = character(),
                      p.value = numeric())
for (e in seq_along(exp)) {
  samples1 <- samples %>% filter(experiment == exp[e])
  controls1 <- controls %>% filter(experiment == exp[e])
  for (a in seq_along(as)) {
    samples2 <- samples1 %>% filter(assay == as[a])
    for (o in seq_along(ods)) {
      samples3 <- samples2 %>% filter(od == ods[o])
      if(nrow(samples3) == 0) next
      test <- t.test(x = samples3$conc, y = controls1$conc, alternative = "two.sided")
      p_value[n, 1:4] <- c(exp[e], as[a], ods[o], test$p.value)
      n <- n + 1
    }
  }
}

write.table(p_value,
            file = "results/fu6_p_value_sn.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)
```


Results summary:
```{r eval = FALSE}
median <- data %>%
  group_by(experiment, assay, od, strain.id, strain, group, compound, unit) %>%
  summarise(conc = median(conc),
            percent = median(percent))

write.table(median,
            file = "results/fu6_results_summary.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```

PFNA is significantly reduced for all ODs and all cultures.


