---
title: "PFNA accumulation range"
author: "Anna Lindell"
date: "20/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(tidyverse)
library(readr)
```



# Quality control

Read in data:
```{r eval = FALSE}
data <- read.table(file = "data/fu7_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)
```


Internal standard:
```{r eval = FALSE}
is <- data %>% filter(group != "Blank")

ggplot(is, aes(x = file, y = is.resp)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_rect(fill="white")) +
  labs(y = "Internal Standard Response", x = "Data File") +
  facet_wrap(~compound, scales = "free_x") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/fu7_is.png",
       width = 12, height = 7, units = "cm")

print.data.frame(is %>% group_by(compound) %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), is.resp))
```

Ibuprofen: Coefficient of variation 
PFNA: 38 %
PFOA: 8.8 %
-> slight signal drift for both compounds
-> for PFNA concentration curve much higher is response



Quality control for response:
```{r eval = FALSE}
qc <- data %>% filter(group == "QC")

ggplot(qc, aes(x = file, y = compound.resp)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_rect(fill="white")) +
  labs(y = "Response", x = "Data File") +
  facet_wrap(~compound, scales = "free_x") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/fu7_qc.png",
       width = 7, height = 5, units = "cm")

print.data.frame(qc %>% group_by(compound) %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), compound.resp))
```

QC: Coefficient of variation 
PFNA: 30.5 %
PFOA: 4.2 %
-> slight signal drift


--> since signal drift in the same direction for Ibuprofen and PFNA ratio used for concentration calculations.


Quality control for ratio:
```{r eval = FALSE}
ggplot(qc, aes(x = file, y = ratio)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_rect(fill="white")) +
  labs(y = "Ratio", x = "Data File") +
  facet_wrap(~compound, scales = "free_x") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/fu7_qc_ratio.png",
       width = 7, height = 5, units = "cm")

print.data.frame(qc %>% group_by(compound) %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), ratio))

rm(list = ls())
```

QC: Coefficient of variation improved
PFNA: 8.3 %
PFOA: 5.7 %





# Median Reduction:

```{r eval = FALSE}
data <- read.table(file = "data/fu7_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data <- data %>%
  filter(group %in% c("Compound control","Supernatant","Whole culture")) 


cc_median <- data %>%
  filter(group == "Compound control") %>%
  group_by(compound, set_conc) %>%
  summarise(median = median(conc))

cc_median <- cc_median$median
cc_median <- rep(cc_median,  each = 9)

median <- data %>%
  arrange(experiment, compound, set_conc, group, replicate) %>%
  mutate(percent = conc / cc_median * 100)

write.table(median,
            file = "data/fu7_medianreduction.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```




# Plot samples

cc, sn, wc:
```{r eval = FALSE}
data <- read.table(file = "data/fu7_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

ggplot(data, aes(x = group, y = percent)) +
  geom_jitter(width = 0.1, size = 3, shape = 1) +
  facet_grid2(rows = vars(compound), cols = vars(set_conc), scales = "free_x", render_empty = FALSE) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 0.5, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        strip.background = element_rect(colour="white", fill="white")) +
  labs(y = "PFAS recovery [%]") +
  ylim(0,NA)
ggsave(file = "results/fu7_percent.pdf",
           width = 20, height = 9, units = "cm")
```


Plot sn depletion:
```{r eval = FALSE}
sn <- data %>% filter(group == "Supernatant")

ggplot(sn, aes(x = percent, y = set_conc)) +
  geom_point(size = 3, shape = 1) +
  theme_bw() +
  xlim(0,NA) +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        strip.background = element_blank()) +
  facet_grid(cols = vars(compound)) +
  labs(x = "Supernatant Recovery [%]", y = "PFAS concentration [uM]")
ggsave(file = "results/fu7_percent_sn.pdf",
           width = 12, height = 7, units = "cm")

rm(list = ls())
```




# Statistical analyis

T-test to test if supernatant significantly reduced:
```{r eval = FALSE}
data <- read.table(file = "data/fu7_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

samples <- data %>% filter(group == c("Supernatant"))
controls <- data %>% filter(group == "Compound control")

comp <- unique(samples$compound)
sconc <- unique(samples$set_conc)

n <- 1
p_value <- data.frame(compound = character(),
                      set_conc = character(),
                      p.value = numeric())
for (c in seq_along(comp)) {
  samples1 <- samples %>% filter(compound == comp[c])
  controls1 <- controls %>% filter(compound == comp[c])
  for (s in seq_along(sconc)) {
    samples2 <- samples1 %>% filter(set_conc == sconc[s])
    controls2 <- controls1 %>% filter(set_conc == sconc[s])
    test <- t.test(x = samples2$conc, y = controls2$conc, alternative = "two.sided")
    p_value[n, 1:3] <- c(comp[c], sconc[s], test$p.value)
    n <- n + 1
  }
}

adj_data <- p_value %>%
  group_by(compound) %>%
  mutate(p_adj = p.adjust(p.value, method = "fdr")) %>%
  arrange(compound, set_conc)

write.table(adj_data,
            file = "results/fu7_p_value_sn.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)
```


Results summary:
```{r eval = FALSE}
median <- data %>%
  group_by(experiment, compound, set_conc, strain.id, strain, group, unit) %>%
  summarise(conc = median(conc),
            percent = median(percent))

write.table(median,
            file = "results/fu7_results_summary.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```





