---
title: "PFNA-dose-different-media"
author: "Anna Lindell"
date: "20/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(tidyverse)
library(readr)
library(ggplot2)
```






# Quality control

Read in data:
```{r eval = FALSE}
data <- read.table(file = "data/fu8_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)
```



Internal standard:
```{r eval = FALSE}
is <- data %>% filter(group != "Blank", group != "QC")

ggplot(is, aes(x = file, y = is.resp)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_blank()) +
  labs(y = "Internal Standard Response", x = "Data File") +
  facet_grid(cols = vars(media), rows = vars(injection.vol), scales = "free") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/fu8_is.png",
       width = 15, height = 10, units = "cm")

print.data.frame(is %>% group_by(injection.vol, media) %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), is.resp))
```

Weirdly the supernatant samples have higher IS levels compared to whole culture and compound control  samples.
Coefficient of variation for caffeine 50 - 60 %




Quality control for response:
```{r eval = FALSE}
qc <- data %>% filter(group == "QC")

ggplot(qc, aes(x = file, y = compound.resp)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_blank()) +
  labs(y = "Response", x = "Data File") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/fu8_qc_resp.png",
       width = 7, height = 5, units = "cm")

print.data.frame(qc %>% 
  summarise_each(funs(median, mean, sd, cv=sd/mean), compound.resp))
```

QC: Coefficient of variation 
7 %
-> At the beginning of the run up and down of signal, then constant.



Quality control for ratio:
```{r eval = FALSE}
ggplot(qc, aes(x = file, y = ratio)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_blank()) +
  labs(y = "Ratio", x = "Data File") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/fu8_qc_ratio.png",
       width = 7, height = 5, units = "cm")

print.data.frame(qc %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), ratio))

rm(list = ls())
```

QC: Coefficient of variation slightly worse: 10 %
---> raw response used for concentration calculations




# Median Reduction:

```{r eval = FALSE}
data <- read.table(file = "data/fu8_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data <- data %>%
  filter(group %in% c("Compound control","Supernatant","Whole culture")) %>%
  na.omit()


cc_median <- data %>%
  filter(group == "Compound control") %>%
  group_by(media, set_conc) %>%
  summarise(median = median(conc)) 

cc_median <- cc_median$median
cc_median <- rep(cc_median,  each = 12)

median <- data %>%
  arrange(experiment, media, set_conc, group, replicate) %>%
  mutate(percent = conc / cc_median * 100)

write.table(median,
            file = "data/fu8_medianreduction.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```




# Plot samples

cc, sn, wc:
```{r eval = FALSE}
data <- read.table(file = "data/fu8_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

ggplot(data, aes(x = group, y = percent)) +
  geom_jitter(width = 0.1, size = 2, shape = 1) +
  facet_grid2(rows = vars(media), cols = vars(set_conc), scales = "free_x", render_empty = FALSE) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=10, angle = 90, vjust = 0.5, hjust = 0.5, color = "black"),
        axis.text.y = element_text(size=10, color = "black"),
        strip.background = element_blank()) +
  labs(y = "PFAS recovery [%]") +
  ylim(0,NA)
ggsave(file = "results/fu8_percent.pdf",
           width = 10, height = 11, units = "cm")
```


Plot sn depletion:
```{r eval = FALSE}
sn <- data %>% filter(group == "Supernatant")
sn$set_conc <- as.factor(sn$set_conc)
levels(sn$set_conc)
sn$set_conc <- factor(sn$set_conc, levels=c("0.001", "0.01", "0.1", "1", "10", "100"))

ggplot(sn, aes(x = percent, y = set_conc)) +
  geom_point(size = 3, shape = 1) +
  theme_bw() +
  xlim(0,NA) +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        strip.background = element_blank()) +
  facet_grid(cols = vars(media)) +
  labs(x = "Supernatant Recovery [%]", y = "PFAS concentration [uM]")
ggsave(file = "results/fu8_percent_sn.png",
           width = 15, height = 7, units = "cm")
```


Plot sn depletion for mGAM only:
```{r eval = FALSE}
mgam <- sn %>% filter(media == "mGAM")

ggplot(mgam, aes(x = percent, y = set_conc)) +
  geom_point(size = 3, shape = 1) +
  theme_bw() +
  xlim(0,100) +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        strip.background = element_blank()) +
  labs(x = "Supernatant Recovery [%]", y = "PFAS concentration [uM]")
ggsave(file = "results/fu8_percent_sn_mgam.pdf",
           width = 7, height = 7, units = "cm")

rm(list = ls())
```



# Plot samples after OD correction

Plot sn depletion:
```{r eval = FALSE}
sn <- read.table(file = "data/fu8_od_correction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

sn$set_conc <- as.factor(sn$set_conc)
levels(sn$set_conc)
sn$set_conc <- factor(sn$set_conc, levels=c("0.01", "0.1", "1", "10", "100"))

ggplot(sn, aes(x = accumulation.per.od, y = set_conc)) +
  geom_point(size = 3, shape = 1) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"),
        strip.background = element_blank()) +
  facet_grid(cols = vars(media), scales = "free_x") +
  labs(x = "Accumulation per OD unit", y = "PFAS concentration [uM]")
ggsave(file = "results/fu8_percent_sn_od_corrected.png",
           width = 15, height = 7, units = "cm")

rm(list = ls())
```

Average OD:
BHI: 0.1
GAM: 0.5
mGAM: 0.95

CAVEAT: bacteria barely grew in BHI, making the results for the OD corrected accumulation questionable.




# Statistical analyis

T-test to test if supernatant significantly reduced:
```{r eval = FALSE}
data <- read.table(file = "data/fu8_medianreduction.txt", 
                   sep = "\t", dec = ".", header = TRUE)

samples <- data %>% filter(group == c("Supernatant"))
controls <- data %>% filter(group == "Compound control")

med <- unique(samples$media)
sconc <- unique(samples$set_conc)

n <- 1
p_value <- data.frame(media = character(),
                      set_conc = character(),
                      p.value = numeric())
for (c in seq_along(med)) {
  samples1 <- samples %>% filter(media == med[c])
  controls1 <- controls %>% filter(media == med[c])
  for (s in seq_along(sconc)) {
    samples2 <- samples1 %>% filter(set_conc == sconc[s])
    controls2 <- controls1 %>% filter(set_conc == sconc[s])
    test <- t.test(x = samples2$conc, y = controls2$conc, alternative = "two.sided")
    p_value[n, 1:3] <- c(med[c], sconc[s], test$p.value)
    n <- n + 1
  }
}

adj_data <- p_value %>%
  group_by(media) %>%
  mutate(p_adj = p.adjust(p.value, method = "fdr")) %>%
  arrange(media, set_conc)

write.table(adj_data,
            file = "results/fu8_p_value_sn.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)
```



Results summary:
```{r eval = FALSE}
median <- data %>%
  group_by(experiment, compound, media, set_conc, strain.id, strain, group, unit) %>%
  summarise(conc = median(conc),
            percent = median(percent))

write.table(median,
            file = "results/fu8_results_summary.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```





