---
title: "pfa_dose-response-growth_different-strains"
author: "Anna Lindell"
date: "27/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
```




# Normalising data by minimum value

Read in data:
```{r  eval = FALSE}
data <- read.table(file = "data/fu9_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)
```


Define function:
```{r  eval = FALSE}
SubtractMin <- function(y) {
  new_data <- y %>%
    group_by(strain, compound, concentration, replicate) %>%
    summarize(min = min(OD))
  new_data <- merge(y, new_data, by = c("strain", "compound", "concentration", "replicate"))
  new_data$OD <- new_data$OD - new_data$min
  df<-data.frame(new_data)
  return(df)
}
```


Normalise data:
```{r  eval = FALSE}
norm <- data %>%
  SubtractMin %>%
  arrange(strain, compound, concentration, replicate, time) %>%
  select(id, strain, compound, concentration, replicate, plate, well, time, OD)
```


Save data:
```{r  eval = FALSE}
write.table(norm, file = "data/fu9_normalised-data.txt", 
            sep = "\t", dec = ".",
            row.names = FALSE, col.names = TRUE, quote = FALSE)
```


Plot normalised data:
```{r  eval = FALSE}
norm$replicate <- as.factor(norm$replicate)
norm$concentration <- as.factor(norm$concentration)
levels(norm$concentration)
norm$concentration <- factor(norm$concentration, levels = c("Control", "500", "250", "125", "62", "31", "15.6", "7.8", "3.9", "2", "1"))
norm$strain<- as.factor(norm$strain)
levels(norm$strain)
norm$strain <- factor(norm$strain, levels = c("B. dorei", "B. thetaiotaomicron", "B. uniformis", "C. aerofaciens", "C. comes", "O. splanchnicus", "P. merdae", "E. coli ED1a", "E. coli IAI1", "E. coli BW25113", "E. coli BW25113 delta-tolC",  "E. coli BW25113 imp4213::FRT", "E. coli C43 (DE3) WT", "E. coli C43 (DE3) delta-ArcA-ArcB-TolC", "E. coli C43 (DE3) delta-ArcA-ArcB"))

comp <- c("PFHpA", "PFOA", "PFNA", "PFDA")

for (i in seq_along(comp)) {
  subset <- norm %>% filter(compound == "DMSO" | compound == comp[i])
  plot <- ggplot(subset, aes(x = time, y = OD, color = replicate)) +
    ggtitle(paste(comp[i], "MIC", sep = " ")) +
    geom_point(size = 0.5, alpha = 0.5) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.y = element_text(color = "black"),
        axis.text.x = element_text(color = "black"),
        strip.background = element_blank()) +
    labs(x = "Time [h]", color = "Replicate") +
    scale_color_manual(values=c("#ff1493","#00ff7f","#3399ff","orange")) +
    facet_grid(rows = vars(strain), cols = vars(concentration), scales = "free")
  ggsave(plot, 
         file = paste("results/growth-curve-by-compound/", comp[i], ".pdf", sep = ""),
         width = 25, height = 35, unit = "cm")
}

strains <- unique(norm$strain)

norm$compound <- as.factor(norm$compound)
levels(norm$compound)
norm$compound <- factor(norm$compound, levels = c("DMSO", "PFHpA", "PFOA", "PFNA", "PFDeA"))

for (i in seq_along(strains)) {
  subset <- norm %>% filter(strain == strains[i])
  plot <- ggplot(subset, aes(x = time, y = OD, color = replicate)) +
    ggtitle(paste(strains[i], "MIC", sep = " ")) +
    geom_point(size = 1) +
    theme_bw() +
    labs(x = "Time [h]", color = "Replicate") +
    scale_color_manual(values=c("#ff1493","#00ff7f","#3399ff","orange")) +
    facet_grid(rows = vars(compound), cols = vars(concentration))
  ggsave(plot, 
         file = paste("results/growth-curve-by-strain/", strains[i], ".png", sep = ""),
         width = 25, height = 15, unit = "cm")
}

rm(list = ls())
```



# Calculating AUC and normalising by control values:

Read in data:
```{r  eval = FALSE}
data <- read.table(file = "data/fu9_normalised-data.txt", 
                   sep = "\t", dec = ".", header = TRUE)
```


Calculate AUC:
```{r  eval = FALSE}
library(bayestestR)

auc_data <- data %>%
  group_by(id, strain, compound, concentration, replicate, plate, well) %>%
  summarize(auc = area_under_curve(time, OD, method = "trapezoid"))
```



Define functions for normalsising data:
```{r eval = FALSE}
Normalisation <- function(y) {
  control <- y %>%
    filter(compound == "DMSO") %>%
    group_by(strain, plate) %>%
    summarize(ControlMedian = median(auc))
  new_data <- merge(y, control, by = c("strain", "plate"))
  new_data$norm_auc <- new_data$auc / new_data$ControlMedian
  df<-data.frame(new_data)
  return(df)
}
```


Normalise data by control median:
```{r eval = FALSE}
auc_data1 <- auc_data %>%
  Normalisation %>%
  dplyr::select(id, strain, compound, concentration, replicate, auc, norm_auc) %>%
  arrange(strain, compound, concentration, replicate)

write.table(auc_data1, file = "data/fu9_auc_data.txt", 
            sep = "\t", dec = ".",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```




# Quality control:

Read in data:
```{r  eval = FALSE}
data <- read.table(file = "data/fu9_auc_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)
```


Normalised data distribution:
```{r eval = FALSE}
# look at normalised data distribution:
ggplot(data, aes(x = norm_auc)) +
  geom_density() +
  geom_vline(xintercept = 1) +
  ggtitle("Distribution of AUC")
ggsave(file = "qualitycontrol/distribution-auc.png",
       width = 20, height = 15, unit = "cm")


# look at normalised data distribution for controls:
controls <- data %>%
  filter(compound == "DMSO")
ggplot(controls, aes(x = norm_auc)) +
    geom_density() +
    ggtitle("Distribution of AUC for Controls")
ggsave(file = "qualitycontrol/distribution-auc-controls.png",
       width = 20, height = 15, unit = "cm")


# look at normalised data distribution for controls grouped by strain:
ggplot(controls, aes(x = norm_auc)) +
    geom_density() +
  facet_wrap(~strain) +
    ggtitle("Distribution of AUC for Controls")
ggsave(file = "qualitycontrol/distribution-auc-controls-strains.png",
       width = 20, height = 15, unit = "cm")

rm(list = ls())
```




# Statistical analysis:

Read in data:
```{r  eval = FALSE}
data <- read.table(file = "data/fu9_auc_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)
```


Z-score normalisation:
```{r eval = FALSE}
control <- data %>%
  filter(compound == "DMSO")
samples <- data %>%
  filter(compound != "DMSO")

strains <- unique(samples$strain)
comp <- unique(samples$compound)
conc <- unique(samples$concentration)
rep <- unique(samples$replicate)


new_data <- data.frame(strain = character(),
                       compound = character(),
                       concentration = character(),
                       replicate = character(),
                       auc = numeric(),
                       norm_auc = numeric(),
                       zscore = numeric())
n <- 1
for (s in seq_along(strains)) {
  subset1 <- samples %>% filter(strain == strains[s])
  control1 <- control %>% filter(strain == strains[s])
  for (c in seq_along(comp)) {
    subset2 <- subset1 %>% filter(compound == comp[c])
    for (i in seq_along(conc)) {
      subset3 <- subset2 %>% filter(concentration == conc[i])
      for (r in seq_along(rep)) {
        subset4 <- subset3 %>% filter(replicate == rep[r])
        zscore <- (subset4$norm_auc - mean(control1$norm_auc)) / sd(control1$norm_auc)
        new_data[n, 1:7] <- c(strains[s],
                              comp[c],
                              conc[i],
                              rep[r],
                              subset4$auc,
                              subset4$norm_auc,
                              zscore)
        n <- n + 1
      }
    }
  }
}

```


Convert z-score to p-value:
```{r eval = FALSE}
new_data <- new_data %>%
  mutate_at(c("replicate", "auc", "norm_auc", "zscore"), as.numeric) %>%
  mutate(p_value = 2 * pnorm(abs(zscore), lower.tail=FALSE))
```


Adjust p-value by number of concentrations:
```{r eval = FALSE}
adj_data <- new_data %>%
  group_by(strain, compound, replicate) %>%
  mutate(p_adj = p.adjust(p_value, method = "fdr")) %>%
  arrange(strain, compound, concentration, replicate)

write.table(adj_data, file = "results/fu9_results.txt",
            sep = "\t", dec = ".",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```


# Plot samples

Create a heatmap by compound:
```{r eval = FALSE}
heatmap <- read.table(file = "results/fu9_results.txt", 
                   sep = "\t", dec = ".", header = TRUE)
heatmap <- heatmap %>%
  group_by(strain, compound, concentration) %>%
  summarize(norm_auc = mean(norm_auc)) %>%
  mutate(inhibition = 100 - (norm_auc * 100))

heatmap$inhibition <- pmax(heatmap$inhibition,0)

heatmap$concentration <- as.factor(heatmap$concentration)
levels(heatmap$concentration)
heatmap$compound <- as.factor(heatmap$compound)
levels(heatmap$compound)
heatmap$compound <- factor(heatmap$compound, levels = c("PFHpA", "PFOA", "PFNA", "PFDA"))
heatmap$strain<- as.factor(heatmap$strain)
levels(heatmap$strain)
heatmap$strain <- factor(heatmap$strain, levels = c("B. dorei", "B. thetaiotaomicron", "B. uniformis", "C. aerofaciens", "C. comes", "O. splanchnicus", "P. merdae", "E. coli ED1a", "E. coli IAI1", "E. coli BW25113", "E. coli BW25113 delta-tolC",  "E. coli BW25113 imp4213::FRT", "E. coli C43 (DE3) WT", "E. coli C43 (DE3) delta-ArcA-ArcB-TolC", "E. coli C43 (DE3) delta-ArcA-ArcB"))

ggplot(heatmap, aes(x = concentration, y = compound)) +
  geom_tile(aes(fill = inhibition), color = "black") +
  geom_text(aes(label = round(inhibition, 0)), size = 2) +
  theme_bw() +
  labs(x = "Concentration [uM]") +
  facet_wrap(~ strain, ncol = 3) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, color = "black"),
        axis.text.y = element_text(color = "black"),
        text = element_text(size = 10),
        strip.background = element_blank()) +
  scale_fill_gradient2(name = "Growth\nInhibition [%]", low = "white",  high = "#36454F", limits=c(0,100))
ggsave(file = "results/fu9_heatmap.png",
       width = 20, height = 15, unit = "cm")

ggplot(heatmap, aes(x = concentration, y = compound)) +
  geom_tile(aes(fill = inhibition), color = "black") +
  theme_bw() +
  labs(x = "Concentration [uM]") +
  facet_wrap(~ strain, ncol = 3) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, color = "black"),
        axis.text.y = element_text(color = "black"),
        text = element_text(size = 10),
        strip.background = element_blank()) +
  scale_fill_gradient2(name = "Growth\nInhibition [%]", low = "white",  high = "#36454F", limits=c(0,100))
ggsave(file = "results/fu9_heatmap_restricted.png",
       width = 20, height = 15, unit = "cm")

write.table(heatmap,
            file = "results/fu9_heatmap.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)
```




Plot data without mutant strains and mutant strains separate:
```{r eval = FALSE}
heatmap1 <- heatmap %>%
  filter(strain %in% c("B. dorei", "B. thetaiotaomicron", "B. uniformis", "C. aerofaciens", "C. comes", "E. coli ED1a", "E. coli IAI1", "O. splanchnicus", "P. merdae"))
heatmap1$strain<- as.factor(heatmap1$strain)
levels(heatmap1$strain)
heatmap1$strain <- factor(heatmap1$strain, levels = c("B. dorei", "B. thetaiotaomicron", "B. uniformis", "C. aerofaciens", "C. comes", "E. coli ED1a", "E. coli IAI1", "O. splanchnicus", "P. merdae"))
heatmap2 <- heatmap %>%
  filter(strain %in% c("E. coli BW25113", "E. coli BW25113 imp4213::FRT", "E. coli BW25113 delta-tolC",  "E. coli C43 (DE3) WT", "E. coli C43 (DE3) delta-ArcA-ArcB", "E. coli C43 (DE3) delta-ArcA-ArcB-TolC"))
heatmap2$strain<- as.factor(heatmap2$strain)
levels(heatmap2$strain)
heatmap2$strain <- factor(heatmap2$strain, levels = c("E. coli BW25113", "E. coli BW25113 imp4213::FRT", "E. coli BW25113 delta-tolC",  "E. coli C43 (DE3) WT", "E. coli C43 (DE3) delta-ArcA-ArcB", "E. coli C43 (DE3) delta-ArcA-ArcB-TolC"))

ggplot(heatmap1, aes(x = concentration, y = compound)) +
  geom_tile(aes(fill = inhibition), color = "black") +
  theme_bw() +
  labs(x = "Concentration [uM]") +
  facet_wrap(~ strain, ncol = 3) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, color = "black"),
        axis.text.y = element_text(color = "black"),
        text = element_text(size = 10),
        strip.background = element_blank()) +
  scale_fill_gradient2(name = "Growth\nInhibition [%]", low = "white",  high = "#6f89a2", limits=c(0,100))
ggsave(file = "results/fu9_heatmap_group1.png",
       width = 20, height = 10, unit = "cm")

ggplot(heatmap2, aes(x = concentration, y = compound)) +
  geom_tile(aes(fill = inhibition), color = "black") +
  theme_bw() +
  labs(x = "Concentration [uM]") +
  facet_wrap(~ strain, ncol = 3) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, color = "black"),
        axis.text.y = element_text(color = "black"),
        text = element_text(size = 10),
        strip.background = element_blank()) +
  scale_fill_gradient2(name = "Growth\nInhibition [%]", low = "white",  high = "#6f89a2", limits=c(0,100))
ggsave(file = "results/fu9_heatmap_group2.png",
       width = 20, height = 7, unit = "cm")

rm(list = ls())
```


Hit calling depending on the number of significant replicates:
```{r eval = FALSE}
# Read in data:
data <- read.table(file = "results/fu9_results.txt", 
                   sep = "\t", dec = ".", header = TRUE)

# filter out hits:
inhibition <- data %>%
  group_by(strain, compound, concentration) %>%
  arrange(strain, compound, concentration) %>%
  filter(p_adj < 0.05 & norm_auc < 0.8)

promotion <- data %>%
  group_by(strain, compound, concentration) %>%
  arrange(strain, compound, concentration) %>%
  filter(p_adj < 0.05 & norm_auc > 1.2)


# Define function for hit calling:
HitCalling <- function(y) {
  Mean_of_hits <- y %>%
    group_by(strain, compound, concentration) %>%
    summarize(auc = mean(auc),
              norm_auc = mean(norm_auc),
              p.value = mean(p_value),
              p.adj = mean(p_adj))
  hit_calling <- y %>%
    dplyr::select(strain, compound, concentration, replicate) %>%
    pivot_wider(id_cols = c(strain, compound, concentration), names_from = replicate, names_prefix = "rep", values_from = replicate) %>%
    relocate(rep1, .after = concentration) %>%
    mutate(rep1 = recode(rep1, "1" = "1", "2" = "1", "3" = "1")) %>%
    mutate(rep2 = recode(rep2, "1" = "1", "2" = "1", "3" = "1")) %>%
    mutate(rep3 = recode(rep3, "1" = "1", "2" = "1", "3" = "1")) %>%
    mutate(rep1 = as.numeric(rep1), rep2 = as.numeric(rep2), rep3 = as.numeric(rep3)) %>%
    mutate(hits = sum(rep1, rep2, rep3, na.rm = TRUE))
  hit_calling <- merge(Mean_of_hits, hit_calling, by = c("strain", "compound", "concentration"))
  hit_calling <- hit_calling %>% arrange(strain, compound, concentration)
  df<-data.frame(hit_calling)
  return(df)
}

hits_inhibition <- HitCalling(inhibition)
hits_inhibition <- hits_inhibition %>% filter(hits >= 2)

write.table(hits_inhibition, file = "results/fu9_hits.txt",
            sep = "\t", dec = ".",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```


Determine MIC
```{r eval = FALSE}
# Read in data:
data <- read.table(file = "results/fu9_hits.txt", 
                   sep = "\t", dec = ".", header = TRUE)

mic <- data %>%
  group_by(strain, compound, concentration) %>%
  arrange(strain, compound, concentration) %>%
  filter(norm_auc < 0.1)

write.table(hits_inhibition, file = "results/fu9_hits.txt",
            sep = "\t", dec = ".",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```




