---
title: "M1: Mouse fecal pfna"
author: "Anna Lindell"
date: "03/06/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(tidyverse)
library(readr)
```




# Quality control

Read in data:
```{r eval = FALSE}
data <- read.table(file = "data/m1_feces_data_raw.txt", 
                   sep = "\t", dec = ".", header = TRUE)
```

Internal standard:
```{r eval = FALSE}
is <- data %>%
  filter(id != "Blank")

ggplot(is, aes(x = file, y = is.resp)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_rect(fill="white")) +
  labs(y = "Internal Standard Response", x = "Data File") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/m1_feces_is.png",
       width = 12, height = 7, units = "cm")

print.data.frame(is %>% group_by(compound) %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), is.resp))
```

27.5 % variation for the internal standard (Caffeine).



Quality control:
```{r eval = FALSE}
qc <- data %>% filter(id == "QC")

ggplot(qc, aes(x = file, y = compound.resp)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_rect(fill="white")) +
  labs(y = "Response", x = "Data File") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/m1_feces_qc_response.png",
       width = 7, height = 5, units = "cm")

ggplot(qc, aes(x = file, y = ratio)) +
  geom_point(size = 1.5) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        text = element_text(size=10),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        strip.background = element_rect(fill="white")) +
  labs(y = "Ratio", x = "Data File") +
  ylim(0, NA)
ggsave(file = "qualitycontrol/m1_feces_qc_ratio.png",
       width = 7, height = 5, units = "cm")

print.data.frame(qc %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), compound.resp))
print.data.frame(qc %>%
  summarise_each(funs(median, mean, sd, cv=sd/mean), ratio))

rm(list = ls())
```

QC: Coefficient of variation 
PFNA response: 25 %
PFNA ratio: 8.6 %




# Plot samples

Read in data:
```{r eval = FALSE}
data <- read.table(file = "data/m1_feces_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)

data$group <- as.factor(data$group)
levels(data$group)
data$group <- factor(data$group, levels = c("Com20 + DMSO", "GF + PFNA", "Com20 + PFNA"))
data$cage <- as.factor(data$cage)

data <- data %>%
  unite(timeloc, c(time.point.d, tissue), sep = " ", remove = FALSE)
```


Plot all samples:
```{r eval = FALSE}
ggplot(data, aes(x = group, y = conc2)) +
  geom_boxplot(width = 0.5, outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 3, shape = 1, aes(color = sex)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        text = element_text(size=12),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 10, color = "black", angle = 90, , vjust = 0.5, hjust = 0.5),
        axis.text.y = element_text(size = 10, color = "black")) +
  facet_wrap(~timeloc, ncol = 6) +
  labs(y = "PFNA [ug/g feces]") +
  scale_y_log10()
ggsave("results/m1_feces_all_dmso_log10.png",
       width = 20, height = 10, units = "cm")

ggplot(data, aes(x = group, y = conc2, color = sex)) +
  geom_jitter(width = 0.1, size = 3, shape = 1) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        text = element_text(size=12),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 10, color = "black", angle = 90, , vjust = 0.5, hjust = 0.5),
        axis.text.y = element_text(size = 10, color = "black")) +
  facet_wrap(~timeloc, ncol = 6) +
  labs(y = "PFNA [ug/g feces]") +
  scale_y_log10()
ggsave("results/m1_feces_all_dmso_log10_sex.png",
       width = 20, height = 10, units = "cm")
```


Plot data without DMSO control:
```{r eval = FALSE}
data1 <- data %>%
  filter(group != "Com20 + DMSO")
ggplot(data1, aes(x = group, y = conc2)) +
  geom_boxplot(width = 0.5, outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 3, shape = 1, aes(color = sex)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        text = element_text(size=12),
        axis.title.x = element_blank(),
        axis.text.x = element_text(color = "black", size = 10, angle = 90, , vjust = 0.5, hjust = 0.5),
        axis.text.y = element_text(color = "black", size = 10)) +
  facet_wrap(~timeloc, ncol = 6) +
  labs(y = "PFNA [ug/g feces]") +
  scale_y_log10()
ggsave("results/m1_feces_all_log10.png",
       width = 20, height = 9, units = "cm")

ggplot(data1, aes(x = group, y = conc2, color = sex)) +
  geom_jitter(width = 0.1, size = 3, shape = 1) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        text = element_text(size=12),
        axis.title.x = element_blank(),
        axis.text.x = element_text(color = "black", size = 10, angle = 90, , vjust = 0.5, hjust = 0.5),
        axis.text.y = element_text(color = "black", size = 10)) +
  facet_wrap(~timeloc, ncol = 6) +
  labs(y = "PFNA [ug/g feces]") +
  scale_y_log10()
ggsave("results/m1_feces_all_log10_sex.png",
       width = 20, height = 10, units = "cm")

rm(list = ls())
```






# Statistical test

Read in data:
```{r eval = FALSE}
data <- read.table(file = "data/m1_feces_data.txt", 
                   sep = "\t", dec = ".", header = TRUE)
```


Check highest PFNA level in DMSO control and set this as cut off level (anything above this will not be considered for testing):
```{r eval = FALSE}
data %>%
  filter(group == "DMSO") %>%
  summarize(min = min(concentration.ugg),
            max = max(concentration.ugg))
```



Check highest PFNA level in GF and Com20:
```{r eval = FALSE}
data %>%
  filter(group != "DMSO") %>%
  group_by(group, time.point) %>%
  summarize(min = min(concentration.ugg),
            max = max(concentration.ugg))
```
Only day 0 samples are lower than the DMSO control max value, so all data points from 3h, 1d, 2d and 3d can be analysed for significance.



Mean PFNA levels
```{r eval = FALSE}
data %>%
  filter(group != "DMSO") %>%
  group_by(group, time.point.d, tissue) %>%
  summarize(mean.pfna = mean(conc2))
```


Remove DMSO controls for testing:
```{r eval = FALSE}
data <- data %>%
  filter(group != "Com20 + DMSO")
```


Do a two sided student t-test:
```{r eval = FALSE}
ttest <- data %>%
  group_by(time.point.d, tissue) %>%
  summarize(p.value = t.test(conc2 ~ group, alternative = "two.sided", paired = FALSE, var.equal =TRUE)$p.value)
```


Correction for multiple testing:
```{r eval = FALSE}
ttest <- ttest %>%
  mutate(p_adj = p.adjust(p.value, method = "fdr"))
write.table(ttest,
            file = "results/m1_feces_pvalues.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)
```
There is a significant difference between GF and Com20 PFNA levels at all tested time points.


Calculate fold changed between medians:
```{r eval = FALSE}
foldchange <- data %>%
  group_by(group, time.point.d, tissue) %>%
  summarize(median = median(conc2)) %>%
  pivot_wider(names_from = group, values_from = median) %>%
  mutate(foldchange = `Com20 + PFNA` / `GF + PFNA`)
write.table(foldchange,
            file = "results/m1_feces_foldchange.txt",
            sep = "\t", dec = ".", na = "NA",
            row.names = FALSE, col.names = TRUE, quote = FALSE)

rm(list = ls())
```























